Legacy Data Confounds Modern Genomics Studies

### 2018-05-11 Draft3.tex
added Mathieu as author and changed sentence about seq tech
A few minor edits and comments

\author[1,2]{Luke Anderson-Trocm\'e}
\author[1,2]{Simon Gravel}

@@

Once mismatch sites were removed from the 1kGP data, the XYZ - W \sgcomment{I would call the effect by the name of the signature rather than harris et al -- more descriptive, less personal.} signal disappears \ref{Figure1} \sgcomment{usually you give a descriptie name to labels, latex will figure out the numbers!}, suggesting a technical artifact rather than a population structure effect. Regressions against different quality metrics provided by the 1kGP revealed that mean mapping quality per mapped base pair was an excellent correlate with prevalence of the XYZ mutational signature in 1kGP, with low-quality individuals showing elevated rates of the signature.  \ref{Figure1}

 individuals carrying these mutations all had lower average quality scores of mapped bases.
This suggested that individuals with lower quality were driving this signal.

@@

To further investigate	 the extent to which low quality individuals were associated to spurious mutations, we performed a genome-wide association (GWA) study  \ref{Figure1}. GWA studies are commonly used to find genomic regions associated to biologically relevant traits. In this case, we are using this analysis in a more unconventional way to identify regions of the genome that are associated to biologically irrelevant traits like SNPs associated to individuals with low quality.

@@

	\subsection{Sequencing quality over time}
We turned our attention to the other populations in the 1kGP, we found that the sequencing done in phase 1 was more variable and overall tended to include lower quality sequencing data \ref{MapQual}.
The sequencing quality of individuals increases over time \ref{MapQual}. 
\todo{What sequencers were used in each phase?}
By 2011 the sequencing quality seems to level off.
\begin{figure}
\includegraphics[width=\hsize,keepaspectratio]{MapQualOverTime.jpg}

>>>>

\author[1,2]{Luke Anderson-Trocm\'e}
\author[1,2]{Mathieu Bourgey}
\author[1,2]{Simon Gravel}

@@

Once mismatch sites were removed from the 1kGP data, the XYZ - W \sgcomment{I would call the effect by the name of the signature rather than harris et al -- more descriptive, less personal.} signal disappears \ref{Figure1} \sgcomment{usually you give a descriptie name to labels, latex will figure out the numbers!}, suggesting a technical artifact rather than a population structure effect. Regressions against different quality metrics provided by the 1kGP revealed that mean mapping quality per mapped base pair was an excellent correlate with prevalence of the XYZ mutational signature in 1kGP, with low-quality individuals showing elevated rates of the signature. (\sgcomment{show supplementary Figure where you did the correlations})

 Thus sequences with low mappability harbor mutations that reproduce poorly across studies and exhibit a particular mutational signature.

@@

To further investigate	 the extent to which low quality individuals \sgcomment{low quality sequence data! "low quality individuals" (= deplorables?) can get you in trouble!}  were associated to spurious mutations, we performed a genome-wide association (GWA) study  \ref{Figure1}. \sgcomment{I would rather have: To identify SNPs that are associated with low quality, we performed }

GWA studies are commonly used to find genomic regions associated to biologically relevant traits. In this case, we are using this analysis in a more unconventional way to identify regions of the genome that are associated to biologically irrelevant traits like SNPs associated to individuals with low quality.

@@

	\subsection{Sequencing quality over time}
We turned our attention to the other populations in the 1kGP, we found that the sequencing done in phase 1 was more variable and overall tended to include lower quality sequencing data \ref{MapQual}.
The sequencing quality of individuals increases over time \ref{MapQual}. 
By 2011 the sequencing quality seems to level off, this also coincides with the phasing out of older sequencing technologies. 
\begin{figure}
\includegraphics[width=\hsize,keepaspectratio]{MapQualOverTime.jpg}
